Primal-dual algorithm for computing best response against the unplanned mixed strategies of an opponent in two-person zero-sum sequential games

### 2014-07-23 paper.tex
typos

Consider a two-person zero-sum sequential game with payoff matrix $A \in \mathbb{R}^{m,n}$. Consider the problem of
finding the best response strategy $x^* \in \mathbb{R}_{+}^n \cap E^{-1}(\{e\})$ for player 1, given a fixed
(possibly mixed) strategy $y_0$ for player 2, where $E \in \{-1, 0, +1\}^{r,n}$ and $(1, 0, 0, ..., 0) =: e \in \mathbb{R}^r$ encode player
1's game plan, consistent with the rules of the game. In the language of convex-optimization, one can readily give the following saddle-point formulation for this problem. Viz,

@@

where $\Delta_{m}^0$ is the singleton $\{y_0\}$ if $y_0$ is known, otherwise it is the $m-$dimensional probability simplex $\Delta_{m} := \{x \in \mathbb{R}_{+}^m \text{ s.t }\textbf{1}^Tx = 1\}$.\\\\
Noting that $i_{E^{-1}(\{e\})}(x) = \underset{\eta \in \mathbb{R}^r}{max}\text{ }\eta^T(e - Ex)$, one can re-write (1) as:
\begin{equation}
  \underset{x \in \mathbb{R}^n}{min}\text{ }\underset{y \in \mathbb{R}^m, \eta \in \mathbb{R}^r}{max}\text{ }[y,\eta]^T\left[ \begin{array}{@{}cc@{}}
    A \\ \hline
    -E
    \end{array} \right]x + G(x) - F^*(y, \eta)
\end{equation}
where:
\begin{equation}
  \begin{aligned}
    G:=i_{\mathbb{R}_{+}^n}\\
    F^*: (y, \eta) \mapsto (i_{\Delta_{m}^0}(y) + e^T\eta)
  \end{aligned}
\end{equation}
with resolvent operators given by
\begin{equation}
  \begin{aligned}
    (1 + \tau \partial G)^{-1}: x \mapsto \Pi_{\mathbb{R}_{+}^n}(x) = (x)_+\\
    (1 + \sigma \partial F^*)^{-1}: (y, \eta) \mapsto (\Pi_{\Delta_{m}^0}(y), \eta - \sigma e)% (z, \eta) \mapsto \bigl(y, \frac{\eta}{1 + \sigma}\bigr)
  \end{aligned}
\end{equation}
This way, we obtain following algorithm which solves (2).

\begin{algorithm}[htb]
\caption{Primal-dual algorithm for computing best response against on plained mixed strategies of opponent}%%  \textbf{Let}
\textbf{Initialize} $\tilde{x}^{(0)} = x^{(0)} \in \mathbb{R}^n$; $y^{(0)} \in \mathbb{R}^{m}$; $\eta^{(0)} \in \mathbb{R}^{m}$;
$\tau, \sigma > 0 \text{ s.t. }\tau\sigma (\|A\|^2 + \|E\|^2) \le 1$; $k = 0$.\\
\Repeat{
  convergence}{
  \begin{eqnarray*}
    y^{(k+1)} &\leftarrow& \Pi_{\Delta_{m}^0}\bigl(y^{(k)} + \sigma A \tilde{x}^{(k)}\bigr)\\
    \eta^{(k+1)} &\leftarrow& \eta^{(k)} - \sigma E \tilde{x}^{(k)} - \sigma e\\
    x^{(k+1)} &\leftarrow& (x^{(k)} - \tau A^Ty^{(k+1)} + E^T\eta^{(k+1)})_+\\
    \tilde{x}^{(k+1)} &\leftarrow& 2x^{(k+1)} - x^{(k)}\\
    k &\leftarrow& k + 1
  \end{eqnarray*}
} \Return $x^{(k)}$
\label{Tab:pseudocode_lbfgs}
\end{algorithm}

>>>>

Consider a two-person zero-sum sequential game with payoff matrix $A \in \mathbb{R}^{m,n}$. Consider the problem of
finding the best response strategy $x^* \in \mathbb{R}_{+}^n \cap E^{-1}(\{e\})$ for player 1, given a fixed
(possibly mixed) strategy $y_0$ for player 2, where $E \in \{-1, 0, +1\}^{r,n}$ and $e \in \mathbb{R}^r$ encode player
1's game plan, consistent with the rules of the game. In the language of convex-optimization, one can readily give the following saddle-point formulation for this problem. Viz,

@@

where $\Delta_{m}^0$ is the singleton $\{y_0\}$ if $y_0$ is known, otherwise it is the $m-$dimensional probability simplex $\Delta_{m} := \{x \in \mathbb{R}_{+}^m \text{ s.t }\textbf{1}^Tx = 1\}$.\\\\
Noting that $i_{E^{-1}(\{e\})}(x) = \underset{\eta \in \mathbb{R}^r}{max}\text{ }\eta^T(e - Ex)$, one can re-write (1) as:
\begin{equation}
  \underset{x \in \mathbb{R}^n}{min}\text{ }\underset{y \in \mathbb{R}^m, \eta \in \mathbb{R}^r}{max}\text{ }[y,\eta]^T\left[ \begin{array}{@{}cc@{}}
    A \\ \hline
    -E
    \end{array} \right]x + G(x) - F^*(y, \eta)
\end{equation}
where:
\begin{equation}
  \begin{aligned}
    G:=i_{\mathbb{R}_{+}^n}\\
    F^*: (y, \eta) \mapsto (i_{\Delta_{m}^0}(y) - e^T\eta)
  \end{aligned}
\end{equation}
with resolvent operators given by
\begin{equation}
  \begin{aligned}
    (1 + \tau \partial G)^{-1}: x \mapsto \Pi_{\mathbb{R}_{+}^n}(x) = (x)_+\\
    (1 + \sigma \partial F^*)^{-1}: (y, \eta) \mapsto (\Pi_{\Delta_{m}^0}(y), \eta + \sigma e)% (z, \eta) \mapsto \bigl(y, \frac{\eta}{1 + \sigma}\bigr)
  \end{aligned}
\end{equation}
This way, we obtain following algorithm which solves (2).

\begin{algorithm}[htb]
\caption{Primal-dual algorithm for computing best response against unplained mixed strategies of opponent}%%  \textbf{Let}
\textbf{Initialize} $\tilde{x}^{(0)} = x^{(0)} \in \mathbb{R}^n$; $y^{(0)} \in \mathbb{R}^{m}$; $\eta^{(0)} \in \mathbb{R}^{m}$;
$\tau, \sigma > 0 \text{ s.t. }\tau\sigma (\|A\|^2 + \|E\|^2) \le 1$; $k = 0$.\\
\Repeat{
  convergence}{
  \begin{eqnarray*}
    y^{(k+1)} &\leftarrow& \Pi_{\Delta_{m}^0}\bigl(y^{(k)} + \sigma A \tilde{x}^{(k)}\bigr)\\
    \eta^{(k+1)} &\leftarrow& \eta^{(k)} - \sigma E \tilde{x}^{(k)} + \sigma e\\
    x^{(k+1)} &\leftarrow& (x^{(k)} - \tau A^Ty^{(k+1)} + E^T\eta^{(k+1)})_+\\
    \tilde{x}^{(k+1)} &\leftarrow& 2x^{(k+1)} - x^{(k)}\\
    k &\leftarrow& k + 1
  \end{eqnarray*}
} \Return $x^{(k)}$
\label{Tab:pseudocode_lbfgs}
\end{algorithm}
